% -*- TeX-master: "main"; fill-column: 72 -*-

\section{Validation of SBML documents}
\label{apdx-validation}

\subsection{Validation and consistency rules}
\label{validation-rules}

This section summarizes all the conditions that must (or in some cases,
at least \emph{should}) be true of an SBML Level~3 Version~1 model that
uses the \SIDReferencesPackage. We use the same conventions as are used
in the SBML Level~3 Version~1 Core specification document. In
particular, there are different degrees of rule strictness. Formally,
the differences are expressed in the statement of a rule: either a rule
states that a condition \emph{must} be true, or a rule states that it
\emph{should} be true. Rules of the former kind are strict SBML
validation rules---a model encoded in SBML must conform to all of them
in order to be considered valid. Rules of the latter kind are
consistency rules. To help highlight these differences, we use the
following three symbols next to the rule numbers:

\begin{description}

\item[\hspace*{6.5pt}\vSymbol\vsp] A \vSymbolName indicates a
\emph{requirement} for SBML conformance. If a model does not follow this
rule, it does not conform to the \SIDReferencesPackage specification.
(Mnemonic intention behind the choice of symbol: ``This must be
checked.'')

\item[\hspace*{6.5pt}\cSymbol\csp] A \cSymbolName indicates a
\emph{recommendation} for model consistency. If a model does not follow
this rule, it is not considered strictly invalid as far as the
\SIDReferencesPackage specification is concerned; however, it indicates
that the model contains a physical or conceptual inconsistency.
(Mnemonic intention behind the choice of symbol: ``This is a cause for
warning.'')

\item[\hspace*{6.5pt}\mSymbol\msp] A \mSymbolName indicates a strong
recommendation for good modeling practice. This rule is not strictly a
matter of SBML encoding, but the recommendation comes from logical
reasoning. As in the previous case, if a model does not follow this
rule, it is not strictly considered an invalid SBML encoding. (Mnemonic
intention behind the choice of symbol: ``You're a star if you heed
this.'')

\end{description}

The validation rules listed in the following subsections are all stated
or implied in the rest of this specification document. They are
enumerated here for convenience. Unless explicitly stated, all
validation rules concern objects and attributes specifically defined in
the \SIDReferencesPackage package.

For \notice convenience and brevity, we use the shorthand
``\token{refs:\-x}'' to stand for an attribute or element name \token{x}
in the namespace for the \SIDReferencesPackage package, using the
namespace prefix \token{refs}. In reality, the prefix string may be
different from the literal ``\token{refs}'' used here (and indeed, it
can be any valid XML namespace prefix that the modeler or software
chooses). We use ``\token{refs:\-x}'' because it is shorter than to
write a full explanation everywhere we refer to an attribute or element
in the \SIDReferencesPackage namespace.

\subsubsection*{General rules about this package}

\validRule{refs-10101}{To conform to the \SIDReferencesPackage
specification for SBML Level~3 Version~1, an SBML document must declare
the use of the following XML Namespace:
\uri{http://www.sbml.org/sbml/level3/version1/refs/version1}.
(Reference: SBML Level~3 Package specification for SID References,
Version~1 \sec{xml-namespace}.)}

\validRule{refs-10102}{Wherever they appear in an SBML document,
elements and attributes from the \SIDReferencesPackage must be declared
either implicitly or explicitly to be in the XML namespace
\uri{http://www.sbml.org/sbml/level3/version1/refs/version1}.
(Reference: SBML Level~3 Package specification for SID References,
Version~1 \sec{xml-namespace}.)}

\subsubsection*{General rules about identifiers}

\validRule{refs-10301}{(Extends validation rule \#10301 in the
\sbmlthreecore specification. TO DO list scope of ids) (Reference: SBML
Level~3 Package specification for SID References, Version~1
\sec{primitive-types}.)}

\validRule{refs-10302}{The value of a \token{refs:\-id} must conform to
the syntax of the \class{SBML} data type \primtype{SId} (Reference: SBML
Level~3 Package specification for SID References, Version~1
\sec{primitive-types}.)}

\TODO{ANY LIST OF ELEMENTS THAT HAVE ATTRIBUTES}

\subsubsection*{Rules for the extended \class{SBML} class}

\validRule{refs-20101}{In all SBML documents using the
\SIDReferencesPackage, the \class{SBML} object must have the
\token{refs:\-required} attribute. (Reference: SBML Level~3 Version~1
Core, Section~4.1.2.)}

\validRule{refs-20102}{The value of attribute \token{refs:\-required} on
the \class{SBML} object must be of data type \primtype{boolean}.
(Reference: SBML Level~3 Version~1 Core, Section~4.1.2.)}

\validRule{refs-20103}{The value of attribute \token{refs:\-required} on
the \class{SBML} object must be set to \val{false}. (Reference: SBML
Level~3 Package specification for SID References, Version~1
\sec{xml-namespace}.)}

\subsubsection*{Rules for extended \class{Model} object}

\validRule{refs-20201}{A \Model object may contain one and only one
instance of each of the \ThingA, \ThingB and \ListOfThingCs elements. No
other elements from the SBML Level 3 SID References namespaces are
permitted on a \Model object. (Reference: SBML Level~3 Package
specification for SID References, Version~1, \sec{model-class}.)}

\validRule{refs-20202}{The \ListOfThingCs subobject on a \Model object
is optional, but if present, this container object must not be empty.
(Reference: SBML Level~3 Specification for SID References Version~1,
\sec{model-class}.)}

\validRule{refs-20203}{Apart from the general notes and annotations
subobjects permitted on all SBML objects, a ListOfThingCs container
object may only contain \ThingC objects. (Reference: SBML Level~3
Package specification for SID References, Version~1,
\sec{model-class}.)}

\validRule{refs-20204}{A \ListOfThingCs object may have the optional
SBML Level~3 Core attributes \token{metaid} and \token{sboTerm}. No
other attributes from the SBML Level 3 Core namespaces are permitted on
a \ListOfThingCs object. (Reference: SBML Level~3 Package specification
for SID References, Version~1, \sec{model-class}.)}


\subsubsection*{Rules for \class{ThingA} object}

\validRule{refs-20301}{A \ThingA object may have the optional SBML
Level~3 Core attributes \token{metaid} and \token{sboTerm}. No other
attributes from the SBML Level 3 Core namespaces are permitted on a
\ThingA. (Reference: SBML Level~3 Version~1 Core, Section~3.2.)}

\validRule{refs-20302}{A \ThingA object may have the optional SBML
Level~3 Core subobjects for notes and annotations. No other elements
from the SBML Level 3 Core namespaces are permitted on a \ThingA.
(Reference: SBML Level~3 Version~1 Core, Section~3.2.)}

\validRule{refs-20303}{A \ThingA object must have the required attribute
\token{refs:\-id}. No other attributes from the SBML Level 3 SID
References namespaces are permitted on a \ThingA object. (Reference:
SBML Level~3 Package specification for SID References, Version~1,
\sec{thinga-class}.)}


\subsubsection*{Rules for \class{ThingB} object}

\validRule{refs-20401}{A \ThingB object may have the optional SBML
Level~3 Core attributes \token{metaid} and \token{sboTerm}. No other
attributes from the SBML Level 3 Core namespaces are permitted on a
\ThingB. (Reference: SBML Level~3 Version~1 Core, Section~3.2.)}

\validRule{refs-20402}{A \ThingB object may have the optional SBML
Level~3 Core subobjects for notes and annotations. No other elements
from the SBML Level 3 Core namespaces are permitted on a \ThingB.
(Reference: SBML Level~3 Version~1 Core, Section~3.2.)}

\validRule{refs-20403}{A \ThingB object must have the required attribute
\token{refs:\-id}. No other attributes from the SBML Level 3 SID
References namespaces are permitted on a \ThingB object. (Reference:
SBML Level~3 Package specification for SID References, Version~1,
\sec{thingb-class}.)}


\subsubsection*{Rules for \class{ThingC} object}

\validRule{refs-20501}{A \ThingC object may have the optional SBML
Level~3 Core attributes \token{metaid} and \token{sboTerm}. No other
attributes from the SBML Level 3 Core namespaces are permitted on a
\ThingC. (Reference: SBML Level~3 Version~1 Core, Section~3.2.)}

\validRule{refs-20502}{A \ThingC object may have the optional SBML
Level~3 Core subobjects for notes and annotations. No other elements
from the SBML Level 3 Core namespaces are permitted on a \ThingC.
(Reference: SBML Level~3 Version~1 Core, Section~3.2.)}

\validRule{refs-20503}{A \ThingC object must have the required
attributes \token{refs:\-thing} and \token{refs:\-id}. No other
attributes from the SBML Level 3 SID References namespaces are permitted
on a \ThingC object. (Reference: SBML Level~3 Package specification for
SID References, Version~1, \sec{thingc-class}.)}

\validRule{refs-20504}{The value of the attribute \token{refs:\-thing}
of a \ThingC object must be the identifier of an existing \ThingA or
\ThingB object defined in the enclosing \Model object. (Reference: SBML
Level~3 Package specification for SID References, Version~1,
\sec{thingc-class}.)}


